---
output:
  pdf_document:
    keep_tex: true
    fig_caption: true
    toc: true
    toc_depth: 3 
    number_sections: true
    citation_package: natbib
    latex_engine: pdflatex
    template: ./../latex-templates/report.tex
  html_document:
    df_print: paged
    
title: "A boring (academic) title or a clever title?"
subtitle: "A secondary title"
author: 
- name: "YOUR NAME HERE"
  affiliation: "Washington State University"
keywords: |
    multiple comparisons to control; multivariate chi-square distribution; nonlinear growth curves; Richard's curve; simulated critical points
abstract: |
  In this article we compare the \emph{empirical characteristic function} \citep{Tukey:1977, Becker:1988} to a \emph{moment-generating-functional form} to compute the proportion of hypotheses $m$ that are rejected under the null hypothesis. \vspace{0.25in}
  \noindent Here is a second paragraph of the abstract (if necessary), and with the pipe notation it doesn't break. Notice it still needs to be indented. \vspace{0.25in}
  \noindent Generally, we write this abstract last.  Often it is called the executive summary. It should succinctly summarize the entire document.  You can include references such as this one to the Appendices section \ref{sec:appendix} if necessary.
sectionnumberdepth: 3
titleEndnotes: "ENDNOTES"
titleReferences: "REFERENCES"
columnsReferences: 2
titleTOC:  "TABLE OF CONTENTS"
bibliography: ./../biblio/master.bib
bibliostyle: ./../biblio/ormsv080.bst
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,appendix-setup,message=FALSE}
library(devtools);       # required for source_url
path.humanVerseWSU = "https://raw.githubusercontent.com/MonteShaffer/humanVerseWSU/"
source_url( paste0(path.humanVerseWSU,"master/misc/functions-project-measure.R") );
source_url( paste0(path.humanVerseWSU,"master/humanVerseWSU/R/functions-dataframe.R") );
source_url( paste0(path.humanVerseWSU,"master/humanVerseWSU/R/functions-EDA.R") );
```


```{r,appendix-secret,message=FALSE}
path.project = "C:/_git_/WSU_STATS419_FALL2020/project-measure/";
path.to.secret = "C:/Users/13608/Dropbox/WSU-419/Fall 2020/__student_access__/_SECRET_/"

measure = utils::read.csv( paste0(path.to.secret, "measure-students.txt"), header=TRUE, quote="", sep="|");
```


```{r covert_unit}

# source functions-project-measure.R from my github repository
path.github = "https://raw.githubusercontent.com/minju-lee92/WSU_STATS419_FALL2020/"
source_url( paste0(path.github,"master/functions/functions-project-measure.R") )

# covert inches to cm
measureAscm <-convert.inchestocm(measure)
```


#### Data collapsing

Some people have data for a person's "left" and "right" side of the body. The function merge.left.right() takes measurement from the left or right, if one is NA it returns the other. If they both are available, it returns the mean.

```{r data_collapsing}
# build merged left/right value cols
getOne = c("hand.length", "hand.width", "hand.elbow", "elbow.armpit", "arm.reach", "foot.length", "floor.kneepit", "floor.hip", "floor.armpit")
merged.df <-merge.left.right(measureAscm, getOne)
```

#### Data cleaning
```{r data_cleaning}
# remove NAs, duplicates, create categorical variables... etc 
cleaned.df = prepareMeasureData(merged.df)
```

#### Data creation
There may be a few data features you may want to create.  I have the "arm span" and information about the "armpits" which would enable you to compute the internal "chest width" (from armpit to armpit).  There may be other data you can create in a similar fashion.

#### Data proportions

It is very likely that for each measure row, you would want to create "scaled variables to that person's height", also known as a proportion.

Alternatively, you could scale everything to a person's head height.

Alternatively, you could review lots of different proportions.  I suggested at one point that the foot-size and the "upper arm" (elbow-pit to arm-pit) are the same size (some basic Pythagorean theorem could get you there or close).

There are lots of possibilities, all depends on your interests.  

- Some say the unit of length of a "one foot" that we now decompose into 12 inches was a function of the actual length of the King's foot in England, and would change when a new King was crowned.  

- Another measure of length, the "cubit" is derived from the Latin word for "elbow"

- Galileo Galilei, the famous Italian polymath, literally sold his body parts when he died (quite the entrepreneur).  He had extremely long fingers.  In the museum in Firenze, they have on display a few of the fingers recovered.  Yes, I have seen them <https://www.museogalileo.it/it/>.  Most people miss this museum because they are too busy admiring David's proportions at the nearby Academia Gallery <https://en.wikipedia.org/wiki/David_(Michelangelo)>.

```{r}
# create scaled variables using height
colnum = c(3:7,20:28)
new.colname = c("height/height", "head.height/height", "head.c/height", "arm.span/height", "floor.navel/height","hand.length/height","hand.width.height","hand.elbow/height","elbow.armpit/height","arm.reach/height","foot.length/height","floor.kneepit/height", "floor.hip/height", "floor.armpit/height")

v1.df =build.scale.variables(cleaned.df, colnum, new.colname, cleaned.df$height)

#create scaled variables using head.height
colnum2 = c(3:7,20:28)
new.colname2 = c("height/head.height", "head.height/head.height", "head.c/head.height", "arm.span/head.height", "floor.navel/head.height","hand.length/head.height","hand.width/head.height","hand.elbow/head.height","elbow.armpit/head.height","arm.reach/head.height","foot.length/head.height","floor.kneepit/head.height", "floor.hip/head.height", "floor.armpit/head.height")

v2.df =build.scale.variables(v1.df, colnum2, new.colname2, v1.df$head.height)
v2.df
```

```{r data_selection}
# height, head.height, arm.span, floor.hip, foot.length, hand.length
sample.data <- v2.df[,c(1:4, 6 ,13,20,25,27 ,29:31, 36,37,39,41, 46,48)]
# age over 18
sample.data <- sample.data[sample.data$age >=18,]
summary(sample.data)
# getting rid of outliers.
sample.data <- sample.data[sample.data$arm.span>100&sample.data$hand.length<50&sample.data$floor.hip>50,]
boxplot(sample.data[,c(3:5,7:9)])$out
sample.data <- sample.data[-which(is.na(sample.data$data_collector)), ]
sample.data
```

```{r corr}
sample.cor <- cor(na.omit(sample.data[,c(3:5,7:9)]))
library(corrgram)
corPlot(sample.cor)
```

```{r,setup-test-correlation,message=FALSE}
path.project = "C:/_git_/WSU_STATS419_FALL2020/project-measure/";
path.tables = paste0(path.project,"tables/");
  createDirRecursive(path.tables);
```


```{r,test-correlation,message=FALSE}
file.correlation = paste0(path.tables,"correlation-table.tex");
myData = as.matrix(sample.data[,c(3:5,7:9)]);  # numeric values only, only what will appear in table
# https://www.overleaf.com/read/srzhrcryjpwn
# keepaspectratio of include graphics 
# could scale \input if still too big ...
# https://tex.stackexchange.com/questions/13460/scalebox-knowing-how-much-it-scales#13487
buildLatexCorrelationTable(myData, 
  rotateTable = TRUE,
  width.table = 0.80, # best for given data ... 0.95 when rotateTable = FALSE
                      # 0.60 when rotateTable = TRUE
  myCaption = "Descriptive Statistics and Correlation Analysis",
  myFile = file.correlation,
  myNames = colnames(myData),
  showOnes = "right")
Sys.sleep(2); # in case Knit-PDF doesn't like that I just created the file...
```

\newpage

\input{tables/correlation-table} 

\newpage


I want to explore possible relationships between ethnicity and measurements, and between gender and measurements.
I can possibly compare by race instead of ethnicity.
Difference between child and adults, female and male...

```{r box_plots}

height.m <- sample.data$height[sample.data$my.gender == 'm']
height.f <- sample.data$height[sample.data$my.gender == 'f']
head.height.m <- sample.data$head.height[sample.data$my.gender == 'm']
head.height.f <- sample.data$head.height[sample.data$my.gender == 'f']
arm.span.m <- sample.data$arm.span[sample.data$my.gender == 'm']
arm.span.f <- sample.data$arm.span[sample.data$my.gender == 'f']
floor.hip.m <- sample.data$floor.hip[sample.data$my.gender == 'm']
floor.hip.f <- sample.data$floor.hip[sample.data$my.gender == 'f']
hand.length.m <- sample.data$hand.length[sample.data$my.gender == 'm']
hand.length.f <- sample.data$hand.length[sample.data$my.gender == 'f']
foot.length.m <- sample.data$foot.length[sample.data$my.gender == 'm']
foot.length.f <- sample.data$foot.length[sample.data$my.gender == 'f']


boxplot(height.m, height.f, arm.span.m, arm.span.f, floor.hip.m, floor.hip.f, 
main = "Body measurements boxplot by gender",
at=c(1,2,4,5,7,8),
xlab = "male vs female",
ylab = "unit(cm)",
names=c("height", "height", "armspan", "armspan", "leg", "leg"),
col = c("orange", 'red'),
border = "brown",
notch = TRUE)

boxplot(head.height.m, head.height.f,hand.length.m, hand.length.f, foot.length.m, foot.length.f,
main = "Body measurements boxplot by gender",
at=c(1,2,4,5,7,8),
xlab = "unit(cm)",
col = c("orange", 'red'),
border = "brown",
notch = TRUE)


# boxplot(height.m, height.f,
# main = "Boxplot of height comparision by gender",
# names=c("male", "female"),
# xlab = "Height(cm)",
# col = c("orange", 'red'),
# border = "brown",
# notch = TRUE)

##############################
# boxplot(height~my.ethnicity,
# data = sample.data,
# main = "Boxplot of height comparision by ethnicity",
# xlab = "Height(cm)",
# col = "orange",
# border = "brown",
# notch = TRUE)

#############################
filtered.measure <-sample.data[sample.data$my.ethnicity=='a' | sample.data$my.ethnicity=='w',]
filtered.measure$my.ethnicity <-factor(filtered.measure$my.ethnicity)

boxplot(height~my.ethnicity,
data = filtered.measure,
main = "Boxplot of height comparison by ethnicity",
names=c("asian","white"),
xlab = "Height(cm)",
col = "orange",
border = "brown",
notch = TRUE)

boxplot(floor.hip~my.ethnicity,
data = filtered.measure,
main = "Boxplot of leg length comparision by ethnicity",
names=c("asian","white"),
xlab = "Height(cm)",
col = "orange",
border = "brown",
notch = TRUE)

a.height.m <- filtered.measure$height[filtered.measure$my.gender == 'm' & filtered.measure$my.ethnicity =='a']
a.height.f <- filtered.measure$height[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='a']
w.height.m <- filtered.measure$height[filtered.measure$my.gender == 'm'& filtered.measure$my.ethnicity =='w']
w.height.f <- filtered.measure$height[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='w']

boxplot(a.height.m, a.height.f, w.height.m, w.height.f,
main = "Boxplot of height comparision by gender",
at=c(1,2,4,5),
names=c("asian male", "asian female", "white male", "white female"),
xlab = "Height(cm)",
col = c("orange", 'red'),
border = "brown",
notch = TRUE)

a.height.m <- filtered.measure$height[filtered.measure$my.gender == 'm' & filtered.measure$my.ethnicity =='a']
a.height.f <- filtered.measure$height[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='a']
w.height.m <- filtered.measure$height[filtered.measure$my.gender == 'm'& filtered.measure$my.ethnicity =='w']
w.height.f <- filtered.measure$height[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='w']

a.armspan.m <- filtered.measure$arm.span[filtered.measure$my.gender == 'm' & filtered.measure$my.ethnicity =='a']
a.armspan.f <- filtered.measure$arm.span[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='a']
w.armspan.m <- filtered.measure$arm.span[filtered.measure$my.gender == 'm'& filtered.measure$my.ethnicity =='w']
w.armspan.f <- filtered.measure$arm.span[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='w']

a.leg.m <- filtered.measure$floor.hip[filtered.measure$my.gender == 'm' & filtered.measure$my.ethnicity =='a']
a.leg.f <- filtered.measure$floor.hip[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='a']
w.leg.m <- filtered.measure$floor.hip[filtered.measure$my.gender == 'm'& filtered.measure$my.ethnicity =='w']
w.leg.f <- filtered.measure$floor.hip[filtered.measure$my.gender == 'f'& filtered.measure$my.ethnicity =='w']

boxplot(a.height.m, w.height.m, a.leg.m, w.leg.m, a.armspan.m , w.armspan.m,
main = "Body measurement comparison between asian male and white male",
at=c(1,2,4,5,7,8),
names=c("A.height","W.height","A.leg","W.leg","A.armspan","W.armspan"),
xlab = "asian male vs white male",
ylab ="unit(cm)",
col = c("orange", 'red'),
border = "brown",
notch = TRUE)
```



```{r gg_boxplots}
library(ggplot2)

ggplot(filtered.measure, aes(x=my.ethnicity, y=height, fill =my.gender)) +
  geom_boxplot(fill = "chocolate", notch = TRUE)+facet_wrap(~ my.gender)
```
```{r proportions}
filtered.measure
ggplot(filtered.measure, aes(x=my.gender, y= floor.hip/height)) +
  geom_boxplot(fill = "chocolate", notch = TRUE)

ggplot(filtered.measure, aes(x=my.gender, y= arm.span/height)) +
  geom_boxplot(fill = "chocolate", notch = TRUE)

ggplot(filtered.measure, aes(x=my.gender, y= foot.length/height)) +
  geom_boxplot(fill = "chocolate", notch = TRUE)


ggplot(filtered.measure, aes(x=my.gender, y= head.height/height)) +
  geom_boxplot(fill = "chocolate", notch = TRUE)
```

